MAESTROeX: A Massively Parallel Low Mach Number Astrophysical Solver

### 2019-06-06 MAESTROeX_I/results.tex
Finished results section

\subsection{White Dwarf Convection}

white dwarf convection runs with 3 algorithms (original, new temporal, new temporal + irregularly-spaced base state)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htb]
\begin{center}
\includegraphics[width=2.75in]{./figs/wdconvect_256_maxT}  \hspace{0.5in}
\includegraphics[width=2.75in]{./figs/wdconvect_256_maxMach}
\caption{\label{fig:wdconvect_128_maxval} (left) Peak temperature, $T_{\text{peak}}$, and (right) peak Mach number 
         in a white dwarf until time of ignition at resolution of $256^3$ for three different MAESTROeX algorithms.}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

To test the performance of MAESTROeX for adaptive, three-dimensional simulations, we make use of the AMReX framework to track a localized feature over time. Figure \ref{fig:amr_grids} illustrates the initial grid structures with two levels of refinement for both planar and spherical geometries where the grid is refined according to the temperature and density profiles, respectively. For each of the problems we tested, the single-level simulation was run using the original temporal scheme and the adaptive simulations using the original and new temporal algorithms. We want to show that the adaptive simulation can give the similar results as the single-level simulation and in a more computationally efficient manner.

@@

We continue to use the full-star white dwarf problem described in Section \ref{sec:scaling} to test adaptive simulations on spherical geometry. The simulation on a single-level grid of $512^3$ resolution took 26.01 s per time step (again averaged over 10 time steps) using the original temporal scheme, and 22.66 s using the new scheme. On the adaptive grid with two levels of refinement, 8.9\% of the cells (1,493,504 out of $256^3$ cells) are refined at the first level and 5.3\% (7,176,192 out of $512^3$ cells) at the second. The adaptive simulation took 11.55 s and 12.48 s using the orginal and new temporal schemes, respectively. Both simulations are computed to $t=400$ s and we choose to use the peak temperature as the diagnostics to compare the results. Figure \ref{fig:wdconvect_amr_Tmax} shows the evolution of the peak temperature for all three runs and shows that the adaptive simulation gives the same qualitative result as the single-level simulation. We do not expect the curves to match up exactly because the governing equations are highly nonlinear, and slight differences in the solution caused by solver tolerance and discretization error can change the details of the results. Each simulation was run on Cori KNL with 512 processors and 16 threads.

>>>>

\subsection{White Dwarf Convection}\label{sec:whitedwarf}

To explore the accuracy and efficiency of the new temporal algorithm, we now analyze in detail three-dimensional, full-star calculations of convection preceding ignition in a sub-Chandrasekhar mass white dwarf. Again, we refer the reader to \cite{MAESTRO_AMR} and \cite{MAESTRO_convection} for setup details. We implement both uniformly- and irregularly-spaced base state with the new temporal algorithm, while only uniform base state spacing is used in the original algorithm. As in Section 3 of \cite{MAESTRO_AMR}, we choose the peak temperature and peak Mach number as the two diagnostics to compare the simulations. Figure \ref{fig:wdconvect_256_maxvar} shows the evolution of both peak temperature and peak Mach number until time of ignition on a single-level grid with resolution of $256^3$. The simulation using the new temporal scheme with uniformly-spaced base state gives the same qualitative results as the original scheme, and predicts very similar time of ignition ($t=7810$ s compared to $t=7850$ s for original algorithm). The simulation using the new temporal scheme with irregularly-spaced base state displays a slightly different peak temperature behavior during the initial transition period $t<150$ s, which results in the difference between the curves post transition. We strongly suspect that this is a result of using different initial model inputs for the star in which the resolution near the center of the star is much coarser with the irregular spacing than the uniform spacing. Fortunately, the simulation with irregular base state spacing still follows the same trend as with uniform spacing, and the star is shown to ignite at an earlier time $t=6840$ s.

Figure \ref{fig:wdconvect_Tmax} shows the peak temperature evolution over the first 1000 s on two grids of differing resolutions, $256^3$ and $512^3$. As previously suspected, the simulation with irregularly-spaced base state agrees much closer to the results from using uniform spacing as the resolution is increased. This is most likely due to the increased resolution of the initial star models for the irregular-spaced base state that is now able to properly capture the same properties near the center of the star as its uniformly-spaced counterpart. This is especially important when computing the base state pressure from base state density since we enforce a constant base state pressure outside of a prescribed cutoff density radius that could potentially be quite different with the two types of base state spacing.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[htb]
\begin{center}
\includegraphics[width=2.75in]{./figs/wdconvect_256_maxT}  \hspace{0.5in}
\includegraphics[width=2.75in]{./figs/wdconvect_256_maxMach}
\caption{\label{fig:wdconvect_256_maxvar} (left) Peak temperature, $T_{\text{peak}}$, and (right) peak Mach number 
         in a white dwarf until time of ignition at resolution of $256^3$ for three different MAESTROeX algorithms.}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

In terms of efficiency, all three simulations on $256^3$ single-level grid were run on Cori haswell with 64 processors and 8 threads per core. The simulation using the new temporal algorithm took 6.75 s per time step with uniformly-spaced base state, which is 13\% faster than the 7.77 s per time step when using the original scheme, and 9.72 s with irregularly-spaced base state. The 25\% increase in run time of the irregularly-spaced algorithm is mainly caused by the much larger base state array and could be significantly reduced if we consider truncating the base state beyond the cutoff density radius.

@@

We now test the performance of MAESTROeX for adaptive, three-dimensional simulations to track a localized feature over time. Figure \ref{fig:amr_grids} illustrates the initial grid structures with two levels of refinement for both planar and spherical geometries where the grid is refined according to the temperature and density profiles, respectively. For each of the problems we tested, the single-level simulation was run using the original temporal scheme and the adaptive simulations using the original and new temporal algorithms. We want to show that the adaptive simulation can give the similar results as the single-level simulation and in a more computationally efficient manner.

@@

We continue to use the full-star white dwarf problem described in Section \ref{sec:whitedwarf} to test adaptive simulations on spherical geometry. The adaptive grid is refined twice by tagging the density at $\rho < 5\times 10^7$ g cm$^{-3}$ on the first level and $\rho < 10^8$ g cm$^{-3}$ on the second level. The simulation on a single-level grid of $512^3$ resolution took 26.01 s per time step (again averaged over 10 time steps).  The adaptive grid has a resolution of $128^3$ at the coarsest level and an effective resolution of $512^3$ at the finest level. On this grid, 8.9\% of the cells (1,493,504 out of $256^3$ cells) are refined at the first level and 5.3\% (7,176,192 out of $512^3$ cells) at the second. The adaptive simulation took 11.55 s, resulting in more than a factor of 2 in speedup. Both simulations are computed to $t=400$ s and we choose to use the peak temperature as the diagnostics to compare the results. Figure \ref{fig:wdconvect_amr_Tmax} shows the evolution of the peak temperature for all three runs and shows that the adaptive simulation gives the same qualitative result as the single-level simulation. We do not expect the curves to match up exactly because the governing equations are highly nonlinear, and slight differences in the solution caused by solver tolerance and discretization error can change the details of the results. Each simulation was run on Cori KNL with 512 processors and 16 threads per core.
